Mammalian protein-coding genes exhibit widespread beneficial mutations that are not adaptive

### 2024-04-22 manuscript/main-biorxiv-2.tex
SNP ontology enrichment

\title{Mammalian protein-coding genes exhibit widespread beneficial mutations that are not adaptive}
\rhead{\scshape Beneficial yet non-adaptive mutations}

@@

    Across our selection classes ($\SphyDel$, $\SphyNeu$ and $\SphyBen$), one can ultimately estimate the proportion of correct and incorrect predictions, leading to an estimation of \textit{precision} and \textit{recall} (fig.~\ref{fig:method}F-G and section~\ref{subsec:precisison_recall} in Materials \& Methods).
    Across 28 populations of different mammal species, mutations predicted to be deleterious at the phylogenetic scale ($\ProbaPopDel$) were indeed purged at the population scale, with a \textit{precision} in the range of 90--97\% (table~\ref{table:proba} and fig.~\ref{fig:homo-afr-results}D for humans).
    Conversely, a \textit{recall} in the range of 96--100\% implied that mutations found to be deleterious at the population scale were most likely also predicted to be deleterious at the phylogenetic scale (table~\ref{table:proba}).
    Altogether, purifying selection is largely predictable and amino acids with negative fitness across mammals have been effectively purged away in each population.

>>>>

\title{Estimating the proportion of beneficial mutations that are not adaptive in mammals}
\rhead{\scshape Beneficial yet non-adaptive mutations}

@@

    Across our selection classes ($\SphyDel$, $\SphyNeu$ and $\SphyBen$), one can ultimately estimate the proportion of correct and incorrect predictions, leading to an estimation of \textit{precision} and \textit{recall} (fig.~\ref{fig:method}F-G and section~\ref{subsec:precisison_recall} in Materials \& Methods).
    Across 28 populations of different mammal species, mutations predicted to be deleterious at the phylogenetic scale ($\SphyDel$) were indeed purged at the population scale, with a \textit{precision} in the range of 90--97\% (table~\ref{table:proba} and fig.~\ref{fig:homo-afr-results}D for humans).
    Conversely, a \textit{recall} in the range of 96--100\% implied that mutations found to be deleterious at the population scale were most likely also predicted to be deleterious at the phylogenetic scale (table~\ref{table:proba}).
    Altogether, purifying selection is largely predictable and amino acids with negative fitness across mammals have been effectively purged away in each population.

### 2024-04-22 manuscript/main-biorxiv.tex
SNP ontology enrichment

% The submission checklist is available at:
% https://docs.google.com/document/d/1aHCF1on0mHTK2DoroPCLP_xto4xwHnGJCGwaL1Vm2UE/edit?usp=sharing

@@

    %In this study, we seek empirical evidence for beneficial back-mutations and test long-standing related questions, such as how common they are and what the probability is for any beneficial mutation observed in current populations to be a back-mutation.
    %By addressing these questions, we provide novel insight into the process of adaptation and, more broadly, into how natural selection has affected protein-coding DNA sequences throughout evolution.
    \subsection*{Fitness landscape reconstruction}
    The mutation-selection framework permits to link the patterns of substitution along a phylogeny with the underlying fitness landscape~\cite{halpern_evolutionary_1998, mccandlish_modeling_2014}.
    Such mutation-selection models applied to protein-coding DNA sequences allow us to estimate relative fitnesses for all amino acids for each site of the sequence, explicitly assuming that the underlying fitness landscape is stable along the phylogenetic tree~\cite{rodrigue_mechanistic_2010, tamuri_estimating_2012, rodrigue_detecting_2017}.
    Importantly, because mutation-selection models at the phylogenetic scale are based on population-genetics equations, their estimates of selection coefficients are directly interpretable as fitness effects for an individual within a population.
    %These methods account for phylogenetic relationships, the structure of the genetic code and accommodate for underlying mutational biases~\cite{rodrigue_mechanistic_2010, tamuri_estimating_2012, latrille_improved_2022}.

    Theoretically, we can thus use large-scale genomic data to assess whether the fitness estimated at the phylogenetic scale predict the fitness effect at the population scale for both deleterious and beneficial back-mutations.
    The placental mammals represent an excellent study system to perform such analysis.
    Having originated $\sim$102 million years ago, they diversified quickly~\cite{foley_genomic_2023}.
    Additionally, polymorphism data are available for many species~\cite{howe_ensembl_2021}, as are high quality protein-coding DNA alignments across the genome~\cite{ranwez_orthomam_2007, scornavacca_orthomam_2019}.
    By performing our analysis on 14,509 orthologous protein-coding genes across 87 species, we focus on sequences under functional constraints across mammals.
    For each gene, fitting the mutation-selection model to the multi-species sequence alignment, assuming that the underlying fitness landscape is stable along the phylogenetic tree, allows us to obtain relative fitnesses for all amino acids for each site of the alignment (fig.~\ref{fig:method}A).
    From the fitness landscape, we can calculate the scaled selection coefficient ($\Sphy$) for any possible amino acid change at each position.
    According to its $\Sphy$ value, we can classify any mutation as either a deleterious mutation toward a less fit amino acid ($\SphyDel \coloneqq \Sphy < -1$), a nearly-neutral mutation ($\SphyNeu \coloneqq -1 < \Sphy <1$) or a beneficial back-mutation toward a known fitter amino acid ($\SphyBen \coloneqq \Sphy > 1$).
    Having identified which potential DNA changes represent beneficial back-mutations (fig.~\ref{fig:method}B), we retrieved polymorphism data from 28 populations belonging to 6 genera (\textit{Equus}, \textit{Bos}, \textit{Capra}, \textit{Ovis}, \textit{Chlorocebus}, and \textit{Homo}) to assess the presence of beneficial back-mutations at the population scale.
    We focused on mutations currently segregating within populations and substitutions in the terminal branches, and checked if any of these observed changes were beneficial back-mutations (fig.~\ref{fig:method}C-E).
    In this study, by integrating large-scale genomic datasets at both phylogenetic and population scales, we have been able to estimate the number of beneficial back-mutations across the entire exome of the six genera (figs.~\ref{fig:method}F-G).

    \begin{figure*}[!ht]
        \centering
        \includegraphics[width=\textwidth, page=1] {artworks/figure.fitness-landscape}
        \caption{
            (A \& B) For a given codon position of a protein-coding DNA sequence, amino acids (x-axis) have different fitness values (y-axis).
            Under a fluctuating fitness landscape (A), these fitnesses change with time.
            The protein sequence follows the moving target defined by the amino-acid fitnesses. Since substitutions are preferentially accepted if they are in the direction of this target, substitutions are, on average, adaptive.
            At the phylogenetic scale (C), beneficial substitutions are common (positive signs), promoting phenotype diversification across species.
            Under a stable fitness landscape (B), most mutations reaching fixation are either slightly deleterious reaching fixation due to drift or are beneficial back-mutations restoring a more optimal amino acid.
            At the phylogenetic scale (D), deleterious substitutions (negative signs) are often reverted via beneficial back-mutations (positive signs), promoting phenotype stability and preserving well-established biological systems.
            Even though, individually, any back-mutation might have a small beneficial effect on its bearer, we expect beneficial back-mutations to be scattered across the genome and the genome-wide signature of beneficial back-mutations to be detectable and quantifiable.}
        \label{fig:fitness-landscape}
    \end{figure*}

    \begin{figure*}[!ht]
        \centering
        \includegraphics[width=\textwidth, page=1] {artworks/figure.method.proba}
        \caption{
            Selection coefficients at the phylogenetic and population scales.
            At the phylogenetic scale (A), we estimated the amino-acid fitness for each site from protein-coding DNA alignments using mutation-selection codon models.
            For every possible mutation, the difference in amino acid fitness before and after the mutation allows us to compute the selection coefficient at the phylogenetic scale ($\Sphy$).
            Depending on $\Sphy$ (B) mutations can be predicted as deleterious ($\SphyDel$), nearly-neutral ($\SphyNeu$) or beneficial back-mutations ($\SphyBen$) toward a fitter amino acid and repairing existing functions.
            At the population scale, each observed single nucleotide polymorphism (SNP) segregating in the population can also be classified according to its $\Sphy$ value (C).
            Occurrence and frequency in the population of non-synonymous polymorphisms, contrasted to synonymous polymorphisms (deemed neutral) is used to estimate selection coefficients (D-E) at the population scale ($\Spop$), for each class of selection ($\SphyDel$, $\SphyNeu$, $\SphyBen$).
            We can thus assess whether $\Sphy$ predicts $\Spop$ and compute \textit{precision} (F) and \textit{recall} (G) for each class.
            The \textit{recall} value for class $\SphyBen$ is the probability of back-mutations among all beneficial ones (G).
        }
        \label{fig:method}
    \end{figure*}

@@

    \begin{figure*}[!ht]
        \centering
        \includegraphics[width=\textwidth, page=1] {artworks/figure.homo-afr-results}
        \caption{
            (A) Distribution of scaled selection coefficients ($\Sphy$), predicted for all possible mutations away from the ancestral human genome.
            Mutations are divided into three classes of selection: deleterious ($\SphyDel$), nearly-neutral ($\SphyNeu$) and beneficial ($\SphyBen$, supposedly beneficial back-mutations).
            (B) Distribution of scaled selection coefficients ($\Sphy$) for all observed substitutions along the \textit{Homo} branch after the \textit{Homo}-\textit{Pan} split.
            If there are fewer substitutions than expected, this class is thus undergoing purifying selection, as is the case for $\SphyDel$.
            (C) The site-frequency spectrum (SFS) in humans of African descent for a random sample of 16 alleles (means in solid lines and standard deviations in color shades) for each class of selection and for synonymous mutations, supposedly neutral (black). The SFS represents the proportion of mutations (y-axis) with a given number of derived alleles in the population (x-axis).
            At high frequencies, deleterious mutations are underrepresented.
            (D) Proportion of beneficial $\ProbaPopDel$, nearly-neutral $\ProbaPopNeu$, and deleterious mutations $\ProbaPopBen$ estimated at the population scale for each class of selection at the phylogenetic scale. Proportions depicted here are not weighted by their mutational opportunities.
        }
        \label{fig:homo-afr-results}
    \end{figure*}

@@

    This study represents an essential step toward integrating the different evolutionary scales necessary to understand the combined effects of mutation, selection, and drift on genome evolution.
    In particular, we have been able to quantify the proportion of beneficial back-mutations among all beneficial mutations, which has only been achievable by combining genome-wide data from both phylogenetic and population scales.
    At the phylogenetic scale, codon diversity at each site of a protein-coding DNA alignment allows for reconstructing an amino acid fitness landscape, assuming that this landscape is stable along the phylogeny.
    These amino-acid fitness landscapes allow us to predict any mutation’s selection coefficient ($\Sphy$) along a protein-coding sequence.
    We can compare these selective effects to observations at the population level (figs.~S1-3).
    By doing so, we confirmed that mutations predicted to be deleterious ($\SphyDel \coloneqq \Sphy < -1$) are purified away in extant populations.
    Our results concur with previous studies showing that SIFT scores~\cite{ng_sift_2003, vaser_sift_2016}, based on amino acid alignments across species, also inform on the deleterious fitness effects exerted at the population scale~\cite{chen_hunting_2021}.
    However – contrary to SIFT scores – our mutation-selection model is parameterized by a fitness function such that changes are directly interpretable as fitness effects.
    In this regard, an interesting prediction of our model is that back-mutations are beneficial because they revert previous deleterious changes.
    We have tested this hypothesis and found that these back-mutations ($\SphyBen \coloneqq \Sphy > 1 $) are indeed beneficial in extant populations.
    We estimated that between 11 and 81\% of all beneficial mutations in mammalian populations have not been driven by adaptation but instead have reversed deleterious substitutions which have accumulated along the phylogeny.
    More specifically, in three out of four populations analyzed, the percentage of beneficial mutations estimated to be back-mutations falls between 20 and 40\%.
    These results suggest that many beneficial mutations are non-adaptive but rather restore ancestral states of higher fitness.
    Hence, we can correctly estimate the extent of adaptive evolution only if we account for the number of beneficial back-mutations~\cite{keightley_what_2010, rice_evolutionarily_2015}.
    Altogether, we argue that we should dissociate positive selection from adaptive evolution and limit the use of adaptive mutations to those that are associated with adaptation to environmental change as such~\cite{charlesworth_other_2007, mustonen_fitness_2009}.

    %We have to acknowledge that the mammalian case study might also not be representative of other clades.
    %Indeed, long-term population sizes are relatively small in mammals, which allow large amounts of deleterious mutations to be fixed, and thus, a lot of opportunities for beneficial back-mutations to arise.
    %It would thus be of great interest to reproduce our analysis with other clades, such as \textit{Drosophila} or \textit{Saccharomyces} which have higher effective population sizes.
    %Indeed when the effective population size is higher, the number of effectively neutral mutational opportunities should be reduced.
    %Allowing for changes in effective population sizes along the phylogeny have been developed~\cite{latrille_inferring_2021}, but are, for the moment, too computationally intensive to be performed genome-wide.
    %This reduction should mechanistically increase the proportion of adaptive substitutions.
    %However, this does not imply that the rate of adaptation (i.e the number of changes in the fitness landscape per site per generation) is higher~\cite{galtier_adaptive_2016}.

    %Regarding the distribution of fitness effects (DFE) estimated at the population scale, there is no consensus on its expected shape~\cite{welch_divergence_2008, bataillon_effects_2014}, although it has been found to be reliably constant between closely related species~\cite{castellano_comparison_2019}.
    %We modelled the DFE as a continuous distribution, but to check for robustness in our results despite the uncertainty of the underlying DFE, we also performed our analysis assuming a discretized DFE.
    %Our results are consistent, with the proportion of back-mutations among all beneficial mutations ranging from XX to XX\% across populations when modelling a DFE as a discrete distributio.
    %Consequently, even though our general framework contrasting the phylogenetic and population scales is theoretically consistent, we acknowledge that estimating the proportion of back-mutations among all beneficial ones is dependent on the the underlying DFE at the population scale and the threshold used to define a beneficial mutation.
    %As a result, further work is required to investigate the shape of the DFE at the population scale.
    %Notably, the reconstructed distribution of $\Sphy$ values obtained in our study has a negative mode (around $\Sphy=-5$), consistent with experimental DFEs obtained by mutation accumulation.
    %Contrarily, continuous DFEs estimated at the population scale have by definition a mode at 0.
    %We argue that methods estimating DFE at the population scale should assume a mode shifted toward negative values of $\Spop$.

    %\subsection*{The role of epistasis in mammalian protein-coding genes}
    %
    %In our study, we focused on mutations toward a more optimal amino-acid at a given site, assuming that amino-acid fitness landscapes are site-specific and also independent of one another.
    %Contrarily, under epistasis, the fitness effect of a mutation at a particular site depends on the amino-acids present at other sites.
    %Epistasis therefore allows for compensatory mutations, which restore fitness through mutations at loci different from where deleterious mutations take place, and this process also leaves signals interpretable as non-adaptive beneficial mutations~\cite{hartl_compensatory_1996, pollock_strong_2014, starr_epistasis_2016}.
    %Also, epistasis has been shown to play a role in the evolution of protein-coding genes, with amino-acid residues in contact within a protein or between proteins tending to co-evolve~\cite{breen_epistasis_2012, starr_epistasis_2016}.
    %Particularly, the residues in contact co-evolve to become more compatible with each other generating an entrenchment, making empirical evidence for beneficial back-mutations harder to find since restoring mutations becomes less and less likely to occur over time~\cite{goldstein_nonadaptive_2015, goldstein_sequence_2017, park_epistatic_2022}.
    %Nevertheless, the fact that we observe such a high proportion of beneficial back-mutations suggests that the underlying assumptions of our model, namely site-independence, implying no epistasis, and a static fitness landscape, are a reasonable approximation for the underlying fitness landscape of proteins.
    %Our results imply that the fitness effects of new mutations are mostly conserved across mammalian orthologs, in agreement with other studies showing that for conserved orthologs with similar structures and functions, models without epistasis provide a reasonable estimate of fitness effects in protein coding genes~\cite{ashenberg_mutational_2013, youssef_consequences_2020}.

%TC:ignore
    \section*{Acknowledgements}
    \label{sec:acknowledgment}
    We gratefully acknowledge the help of Mélodie Bastian, Nicolas Lartillot, Carina Farah Mugal, Laurent Duret, Alexandre Reymond, Daniele Silvestro and Nicolas Gambardella for their advice and reviews concerning this manuscript.
    This work was performed using the computing facilities of the CC LBBE/PRABI\@.
    This study makes use of data generated by the NextGen Consortium.
    \textbf{Funding:}
    Université de Lausanne; Agence Nationale de la Recherche, Grant ANR-19-CE12-0019 / HotRec.
    The European Union’s Seventh Framework Programme (FP7/2010-2014) provided funding for the project under grant agreement no 244356 - “NextGen”.
    \textbf{Author contributions:}
    Original idea: T.L.\ and J.J.;
    Model conception: T.L., J.J.\ and N.S.;
    Code: T.L.;
    Data analyses: T.L.\ and J.J.;
    Interpretation: T.L., J.J., D.A.H.\ and N.S.;
    First draft: T.L.\ and J.J.;
    Editing and revisions: T.L., J.J., D.A.H.\ and N.S.
    Project management and funding: N.S\@.
    \textbf{Competing interests:}
    The authors declare no conflicts of interest.
    \textbf{Data and materials availability:}
    The data underlying this article are available at \url{https://doi.org/10.5281/zenodo.7878954}.
    Snakemake pipeline, analysis scripts and documentation are available at \href{https://github.com/ThibaultLatrille/SelCoeff}{github.com/ThibaultLatrille/SelCoeff}.

    \begin{equation}
        \proba [ \SpopBen ] = \sum_{\Sphyclass \in \{\SphyDel, \SphyNeu, \SphyBen \} }\proba [\SpopBen \given \Sphyclass ] \times \proba [\Sphyclass ].
        \label{eq:total_proba}
    \end{equation}
    \printbibliography
\end{document}
%TC:endignore

>>>>

% The submission checklist is available at:
% https://docs.google.com/document/d/1SScaqEAgZGpeaXxpWPDh1NOLaDDLLmkGOYgaGvbY-4o/edit#

@@

    \subsection*{Fitness landscape reconstruction}
    The mutation-selection framework permits to link the patterns of substitution along a phylogeny with the underlying fitness landscape~\cite{halpern_evolutionary_1998, mccandlish_modeling_2014}.
    Such mutation-selection models applied to protein-coding DNA sequences allow us to estimate relative fitnesses for all amino acids for each site of the sequence, explicitly assuming that the underlying fitness landscape is stable along the phylogenetic tree~\cite{rodrigue_mechanistic_2010, tamuri_estimating_2012, rodrigue_detecting_2017}.
    Importantly, because mutation-selection models at the phylogenetic scale are based on population-genetics equations, their estimates of selection coefficients are directly interpretable as fitness effects for an individual within a population.

    Theoretically, we can thus use large-scale genomic data to assess whether the fitness estimated at the phylogenetic scale predict the fitness effect at the population scale for both deleterious and beneficial back-mutations.
    The placental mammals represent an excellent study system to perform such analysis.
    Having originated $\sim$102 million years ago, they diversified quickly~\cite{foley_genomic_2023}.
    Additionally, polymorphism data are available for many species~\cite{howe_ensembl_2021}, as are high quality protein-coding DNA alignments across the genome~\cite{ranwez_orthomam_2007, scornavacca_orthomam_2019}.Mutations can be beneficial by bringing innovation to their bearer, allowing them to adapt to environmental change.
        However, mutations can also be beneficial because they are reverting previous deleterious changes, simply restoring pre-existing functions.
        By integrating phylogenomic and population data in mammals, we estimated the contribution of beneficial back-mutations to molecular evolution, which had remained widely overlooked so far.
        Applying a mutation-selection model, we estimated amino acid fitness landscapes for mammalian protein-coding genes.
        Assuming these fitness landscapes do not change, any beneficial mutation is necessarily a back-mutation.
        We confirmed that these beneficial back-mutations are positively selected in extant populations and demonstrated that a substantial part of ongoing positive selection is not driven by adaptation to environmental change.
    By performing our analysis on 14,509 orthologous protein-coding genes across 87 species, we focus on sequences under functional constraints across mammals.
    For each gene, fitting the mutation-selection model to the multi-species sequence alignment, assuming that the underlying fitness landscape is stable along the phylogenetic tree, allows us to obtain relative fitnesses for all amino acids for each site of the alignment (fig.~\ref{fig:method}A).
    From the fitness landscape, we can calculate the scaled selection coefficient ($\Sphy$) for any possible amino acid change at each position.
    According to its $\Sphy$ value, we can classify any mutation as either a deleterious mutation toward a less fit amino acid ($\SphyDel \coloneqq \Sphy < -1$), a nearly-neutral mutation ($\SphyNeu \coloneqq -1 < \Sphy <1$) or a beneficial back-mutation toward a known fitter amino acid ($\SphyBen \coloneqq \Sphy > 1$).
    Having identified which potential DNA changes represent beneficial back-mutations (fig.~\ref{fig:method}B), we retrieved polymorphism data from 28 populations belonging to 6 genera (\textit{Equus}, \textit{Bos}, \textit{Capra}, \textit{Ovis}, \textit{Chlorocebus}, and \textit{Homo}) to assess the presence of beneficial back-mutations at the population scale.
    We focused on mutations currently segregating within populations and substitutions in the terminal branches, and checked if any of these observed changes were beneficial back-mutations (fig.~\ref{fig:method}C-E).
    In this study, by integrating large-scale genomic datasets at both phylogenetic and population scales, we have been able to estimate the number of beneficial back-mutations across the entire exome of the six genera (figs.~\ref{fig:method}F-G).

    \section*{Results}

@@

    This study represents an essential step toward integrating the different evolutionary scales necessary to understand the combined effects of mutation, selection, and drift on genome evolution.
    In particular, we have been able to quantify the proportion of beneficial back-mutations among all beneficial mutations, which has only been achievable by combining genome-wide data from both phylogenetic and population scales.
    At the phylogenetic scale, codon diversity at each site of a protein-coding DNA alignment allows for reconstructing an amino acid fitness landscape, assuming that this landscape is stable along the phylogeny.
    These amino-acid fitness landscapes allow us to predict any mutation’s selection coefficient ($\Sphy$) along a protein-coding sequence.
    We can compare these selective effects to observations at the population level (figs.~S1-3).
    By doing so, we confirmed that mutations predicted to be deleterious ($\SphyDel \coloneqq \Sphy < -1$) are purified away in extant populations.
    Our results concur with previous studies showing that SIFT scores~\cite{ng_sift_2003, vaser_sift_2016}, based on amino acid alignments across species, also inform on the deleterious fitness effects exerted at the population scale~\cite{chen_hunting_2021}.
    However – contrary to SIFT scores – our mutation-selection model is parameterized by a fitness function such that changes are directly interpretable as fitness effects.
    In this regard, an interesting prediction of our model is that back-mutations are beneficial because they revert previous deleterious changes.
    We have tested this hypothesis and found that these back-mutations ($\SphyBen \coloneqq \Sphy > 1 $) are indeed beneficial in extant populations.
    We estimated that between 11 and 81\% of all beneficial mutations in mammalian populations have not been driven by adaptation but instead have reversed deleterious substitutions which have accumulated along the phylogeny.
    More specifically, in 21 out of 28 populations analyzed, the percentage of beneficial mutations estimated to be back-mutations falls between 20 and 40\%.
    These results suggest that many beneficial mutations are non-adaptive but rather restore ancestral states of higher fitness.
    Hence, we can correctly estimate the extent of adaptive evolution only if we account for the number of beneficial back-mutations~\cite{keightley_what_2010, rice_evolutionarily_2015}.
    Altogether, we argue that we should dissociate positive selection from adaptive evolution and limit the use of adaptive mutations to those that are associated with adaptation to environmental change as such~\cite{charlesworth_other_2007, mustonen_fitness_2009}.

    \begin{equation}
        \proba [ \SpopBen ] = \sum_{\Sphyclass \in \{\SphyDel, \SphyNeu, \SphyBen \} }\proba [\SpopBen \given \Sphyclass ] \times \proba [\Sphyclass ].
        \label{eq:total_proba}
    \end{equation}

    \section*{Data availability}
    The data underlying this article are available at \url{https://doi.org/10.5281/zenodo.7878954}.
    Snakemake pipeline, analysis scripts and documentation are available at \href{https://github.com/ThibaultLatrille/SelCoeff}{github.com/ThibaultLatrille/SelCoeff}.

    \section*{Acknowledgements}
    \label{sec:acknowledgment}
    We gratefully acknowledge the help of Mélodie Bastian, Nicolas Lartillot, Carina Farah Mugal, Laurent Duret, Alexandre Reymond, Daniele Silvestro and Nicolas Gambardella for their advice and reviews concerning this manuscript.
    This work was performed using the computing facilities of the CC LBBE/PRABI\@.
    This study makes use of data generated by the NextGen Consortium.
    Funding:
    Université de Lausanne; Agence Nationale de la Recherche, Grant ANR-19-CE12-0019 / HotRec.
    The European Union’s Seventh Framework Programme (FP7/2010-2014) provided funding for the project under grant agreement no 244356 - “NextGen”.

    \section*{Competing interests}
    The authors declare no conflicts of interest.

    \section*{Author information}
    Original idea: T.L.\ and J.J.;
    Model conception: T.L., J.J.\ and N.S.;
    Code: T.L.;
    Data analyses: T.L.\ and J.J.;
    Interpretation: T.L., J.J., D.A.H.\ and N.S.;
    First draft: T.L.\ and J.J.;
    Editing and revisions: T.L., J.J., D.A.H.\ and N.S.
    Project management and funding: N.S\@.

%TC:ignore
    \printbibliography
%TC:endignore
\newpage
    \begin{figure*}[!htb]
        \centering
        \includegraphics[width=\textwidth, page=1] {artworks/figure.fitness-landscape}
        \caption{
            (A \& B) For a given codon position of a protein-coding DNA sequence, amino acids (x-axis) have different fitness values (y-axis).
            Under a fluctuating fitness landscape (A), these fitnesses change with time.
            The protein sequence follows the moving target defined by the amino-acid fitnesses. Since substitutions are preferentially accepted if they are in the direction of this target, substitutions are, on average, adaptive.
            At the phylogenetic scale (C), beneficial substitutions are common (positive signs), promoting phenotype diversification across species.
            Under a stable fitness landscape (B), most mutations reaching fixation are either slightly deleterious reaching fixation due to drift or are beneficial back-mutations restoring a more optimal amino acid.
            At the phylogenetic scale (D), deleterious substitutions (negative signs) are often reverted via beneficial back-mutations (positive signs), promoting phenotype stability and preserving well-established biological systems.
            Even though, individually, any back-mutation might have a small beneficial effect on its bearer, we expect beneficial back-mutations to be scattered across the genome and the genome-wide signature of beneficial back-mutations to be detectable and quantifiable.}
        \label{fig:fitness-landscape}
    \end{figure*}

    \begin{figure*}[!h]
        \centering
        \includegraphics[width=\textwidth, page=1] {artworks/figure.method.proba}
        \caption{
            Selection coefficients at the phylogenetic and population scales.
            At the phylogenetic scale (A), we estimated the amino-acid fitness for each site from protein-coding DNA alignments using mutation-selection codon models.
            For every possible mutation, the difference in amino acid fitness before and after the mutation allows us to compute the selection coefficient at the phylogenetic scale ($\Sphy$).
            Depending on $\Sphy$ (B) mutations can be predicted as deleterious ($\SphyDel$), nearly-neutral ($\SphyNeu$) or beneficial back-mutations ($\SphyBen$) toward a fitter amino acid and repairing existing functions.
            At the population scale, each observed single nucleotide polymorphism (SNP) segregating in the population can also be classified according to its $\Sphy$ value (C).
            Occurrence and frequency in the population of non-synonymous polymorphisms, contrasted to synonymous polymorphisms (deemed neutral) is used to estimate selection coefficients (D-E) at the population scale ($\Spop$), for each class of selection ($\SphyDel$, $\SphyNeu$, $\SphyBen$).
            We can thus assess whether $\Sphy$ predicts $\Spop$ and compute \textit{precision} (F) and \textit{recall} (G) for each class.
            The \textit{recall} value for class $\SphyBen$ is the probability of back-mutations among all beneficial ones (G).
        }
        \label{fig:method}
    \end{figure*}

    \begin{figure*}[!h]
        \centering
        \includegraphics[width=\textwidth, page=1] {artworks/figure.homo-afr-results}
        \caption{
            (A) Distribution of scaled selection coefficients ($\Sphy$), predicted for all possible mutations away from the ancestral human genome.
            Mutations are divided into three classes of selection: deleterious ($\SphyDel$), nearly-neutral ($\SphyNeu$) and beneficial ($\SphyBen$, supposedly beneficial back-mutations).
            (B) Distribution of scaled selection coefficients ($\Sphy$) for all observed substitutions along the \textit{Homo} branch after the \textit{Homo}-\textit{Pan} split.
            If there are fewer substitutions than expected, this class is thus undergoing purifying selection, as is the case for $\SphyDel$.
            (C) The site-frequency spectrum (SFS) in humans of African descent for a random sample of 16 alleles (means in solid lines and standard deviations in color shades) for each class of selection and for synonymous mutations, supposedly neutral (black). The SFS represents the proportion of mutations (y-axis) with a given number of derived alleles in the population (x-axis).
            At high frequencies, deleterious mutations are underrepresented.
            (D) Proportion of beneficial $\ProbaPopDel$, nearly-neutral $\ProbaPopNeu$, and deleterious mutations $\ProbaPopBen$ estimated at the population scale for each class of selection at the phylogenetic scale. Proportions depicted here are not weighted by their mutational opportunities.
        }
        \label{fig:homo-afr-results}
    \end{figure*}

@@

\end{document}
